---
output: html_document
---

<style>

table, td, th {
  border: none;
  padding-left: 1em;
  padding-right: 1em;
  min-width: 50%;
  margin-left: auto;
  margin-right: auto;
  margin-top: 1em;
  margin-bottom: 1em;
}

</style>

<!-- [![](`r if(dir.exists(add_election_path("images"))) add_election_path("/") else ""`images/icon.png){width=15%} -->
<!-- ![](`r if(dir.exists(add_election_path("images"))) add_election_path("/") else ""`images/logo.png){width=50%}](https://sixtysixwards.com) -->

```{r setup, include=FALSE, out.width='80%'}
# setwd("C:/Users/Jonathan Tannen/Dropbox/sixty_six/posts/turnout_tracker/tracker_v0/")

library(dplyr)
library(stargazer)
library(readr)

source(add_election_path("config.R"), chdir=TRUE)
source("R/util.R", chdir=TRUE)
source("R/generate_plots.R", chdir=TRUE)
source("R/bootstrap.R", chdir=TRUE)

config <- extend_config(config)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)

params <- readRDS(output_file("params.Rds"))
bs <- readRDS(output_file("bootstrap.Rds"))

get_ward <- config$get_ward_from_precinct

raw_data <- readRDS(output_file("raw_data.Rds")) %>% 
  mutate(
    ward=get_ward(precinct),
    time_of_day=config$base_time + minutes(minute)
  )

current_time <- max(bs@raw_result@time_df$time_of_day)

turnout_df <- get_ci_from_bs(bs, predict_topline, keys="time_of_day")
current_turnout <- filter_to_eod(turnout_df)

ward_turnout <- get_ci_from_bs(
  bs, 
  predict_ward_turnout, 
  get_ward=get_ward,
  keys="ward"
)

precinct_turnout <- get_ci_from_bs(
  bs, 
  predict_precinct_eod,
  keys="precinct"
)

write.csv(
  precinct_turnout, 
  file=output_file(sprintf("precinct_turnout_%s.csv", config$city_filename)), 
  row.names=FALSE
)

wards <- readRDS(add_election_path("data/wards.Rds"))
wards <- wards %>% left_join(ward_turnout)

## for text reference
attach(config)

mail_in <- readRDS(output_file("mail_in.Rds"))
mail_count <- get_mail_count(save_dir=OUTPUT_DIR)

save_plot <- function(p, stem){
  file <- sprintf(
    output_file("%s_%s.png"),
    stem,
    format(Sys.time(), "%H_%M_%S")
  )
  ggsave(p, file=file)
}
```
# Live `r substr(election_ds, 1, 4)` Election Turnout Tracker

Welcome to the [`r site_name`](`r site_link`) Turnout Tracker! 
Voters across `r city` are sharing their turnout to support citizen science. 

> First, vote! Then, share your `r precinct_name` and Voter Number at bit.ly/sixtysixturnout .

## Live Results

```{r turnout, echo=FALSE, out.width='100%'}
tp <- turnout_plot(bs, raw_data, config, mail_in=0)
save_plot(tp, "turnout")
tp
```

As of `r pretty_time(current_time)`, `r scales::comma(nrow(raw_data))` voters in `r length(unique(raw_data$ward))` wards have [shared their voter number](`r submission_bitly`). I estimate that <b>`r scales::comma(round(current_turnout$turnout))`</b> people have voted in person across the city (95% Confidence Interval: `r scales::comma(round(current_turnout$p025))` - `r scales::comma(round(current_turnout$p975))`). In addition, more than <b>`r scales::comma(351e3)`</b> mail-in ballots have already been returned*, and `r scales::comma(sum(mail_in$ballots) - 351e3)` are still outstanding.

NOTE: The above mail counts are received from the Commissioners' Office. Some mail-ins that have already arrived are probably not included in this projection. In addition, ballots postmarked today that haven't arrived yet are not included.

## Maps
I use historic correlations among `r precinct_name`s' turnout to estimate turnout in the `r precinct_name`s with no data yet. I simultaneously estimate (1) the overall city-wide turnout, (2) `r precinct_name`s' over-/under-performance, and (3) the city-wide time pattern of voting. See the [Methodology](https://htmlpreview.github.io/?https://github.com/jtannen/turnout_tracker/blob/master/methodology.html).

### Total Turnout (In-Person + Mail-In)
```{r turnout_total_map, echo=FALSE, out.width='100%', fig.height=6, fig.align="center"}
# turnout_map <- gg_wards_predicted(wards, current_time, config)
turnout_map <- leaflet_wards_total(wards, raw_data, mail_in, config, current_time)
turnout_map
```

### Total Turnout (In-Person + Mail-In) vs 2016
```{r turnout_total_change, echo=FALSE, out.width='100%', fig.height=6, fig.align="center"}
# turnout_map <- gg_wards_predicted(wards, current_time, config)
if(sum(mail_in$ballots) > 0){
  change_map <- leaflet_change(wards, raw_data, mail_in, config, current_time)
} else {
  change_map <- ""
}
change_map
```


### In-Person Turnout
```{r turnout_map, echo=FALSE, out.width='100%', fig.height=6, fig.align="center"}
# turnout_map <- gg_wards_predicted(wards, current_time, config)
inperson_map <- leaflet_wards_predicted(wards, raw_data, mail_in, config, current_time)
inperson_map
```

### Mail-In Requests
```{r req_map, echo=FALSE, out.width='100%', fig.height=6, fig.align="center"}
# turnout_map <- gg_wards_predicted(wards, current_time, config)
if(sum(mail_in$ballots) > 0){
  req_map <- leaflet_mail_in_reqs(wards, raw_data, mail_in, config, current_time)
} else {
  req_map <- ""
}
req_map
```

### Mail-In Ballots Returned
```{r ret_map, echo=FALSE, out.width='100%', fig.height=6, fig.align="center"}
# # turnout_map <- gg_wards_predicted(wards, current_time, config)
ret_map <- leaflet_mail_in_returned(wards, raw_data, mail_in, config, current_time)
ret_map
```

### Raw Responses (for debugging)
```{r raw_plot, echo=FALSE, out.width='100%', fig.height=6, fig.align="center"}
turnout_17 <- read.csv(add_election_path(config$turnout_df_path)) %>%
  filter(election=="2017 primary") %>%
  mutate(precinct = sprintf("%04i", precinct))

generate_raw_plot(raw_data, turnout_17, 2017)
```

### Submissions
```{r ward_table, echo=FALSE, out.width='100%', fig.height=6, fig.align="center"}
leaflet_submissions(
  wards=wards, 
  raw_data=raw_data, 
  mail_in=mail_in,
  config=config, 
  current_time=current_time
)
```

## Data Tables
For those of you who like tables with numbers.
```{r map_labelled, echo=FALSE, out.width='100%'}
gg_wards_labelled(wards)
```

```{r table, results='asis'}
stargazer(
  ward_turnout %>% 
    mutate(
      time_of_day=pretty_time(current_time)
    ) %>%
    select(ward, time_of_day, turnout, turnout_p025, turnout_p975) %>%
    mutate_at(
      vars(turnout, turnout_p025, turnout_p975),
      funs(round)
    ),
  type = "html",
  title = "Estimated Turnout by Ward",
  summary = FALSE,
  rownames = FALSE
)
```

## Raw Submissions

```{r raw_data, results='asis'}
stargazer(
  raw_data %>% 
    select(time_of_day, precinct, obs) %>%
    mutate(time_of_day = pretty_time(time_of_day)) %>%
    rename(
      `Time of Vote` = time_of_day, 
      `Precinct`=precinct, 
      `Voter Number` = obs
    ), 
  type = "html",
  title = "Submitted Data", 
  summary = FALSE,
  rownames = FALSE
)
```

**Download the csvs**: - [Raw Submissions](`r "https://jtannen.github.io/raw_data.csv"`) - [Division Predictions](`r sprintf("https://jtannen.github.io/precinct_turnout_%s.csv", config$city_filename)`)